IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-08 paper/intro.tex

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve the language JavaScript, and retrofit dynamic information
flow control on top of it.  A key challenge in dynamic information
flow control is implementation efficiency: when
information flow is tracked at a fine-grained level,
e.g., per value in the system, systems may incur
impractical runtime overhead~\cite{JSFlow}.
%
Moreover, the task of adding fine-grained IFC even to a simple language is
non-trivial~\cite{hritcu2013testing}; when considering a real-world language,
such as JavaScript, this challenge becomes almost insurmountable.

In the alternative \textit{coarse-grained} approach to IFC,
a system is divided into relatively coarse computational units,
each with a single label dictating its security policy.
Only
communication between isolated computational units must be tracked.
A recent system named SWAPI~\cite{swapi} adds IFC to the
browser by identifying \emph{web workers} as an existing computational
unit.  Communication between workers is mediated using
IFC to ensure non-interference, preventing sensitive workers from
influencing the computation of less sensitive workers.
This coarse-grained approach provides a number of advantages:
(1) reduced runtime overhead because checks need only
be performed at isolation boundaries, (2) minimal changes to an
existing programming language when adding IFC, which (3) allows
the reuse of existing programs.  Finally, (4) associating
a single security label with the entire computational unit simplifies
understanding and reasoning about the correctness of the
system, without information flow reasoning about the
technical details of the semantics of the programming language.

A further benefit of the coarse-grained approach
is that a single IFC design and semantics can be generalized
to a large class of programming languages.
Leveraging this insight, we define a general form of coarse-grained
dynamic IFC and show how it can be combined with \emph{any} programming
language conforming to a general semantic framework.
At the core of our IFC system is a sandboxing construct that isolates
potentially sensitive computations, allowing them to
communicate only through clearly specified interfaces.
Using the Matthews-Findler approach~\cite{Matthews:2007:OSM:1190216.1190220}
for combining operational semantics,
we formally state and prove non-interference guarantees that are
independent of the choice of specific target language.

One challenge in the design of our system is that the semantics for the
retrofitted IFC language may not accurately correspond to the optimized
implementation details of a particular target language.
We discuss and partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then non-interference in the abstract specification
carries over to the concrete specification.

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  information flow control language.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Section~\ref{sec:retrofit})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC language that preserves non-interference. (Section~\ref{sec:formal})
  \item We connect our formal semantics to real implementations of
  coarse-grained IFC systems for JavaScript and Haskell, and explain
  how our system could be used to enforce IFC in C. (Section~\ref{sec:real})
  \item We briefly describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, privileges, or a
  notion of clearance. (Section~\ref{sec:extensions})
\end{itemize}

>>>>

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve the language JavaScript, and retrofit dynamic information
flow control on top of it.

Systems that track information flow at a fine-grained level, e.g. per
value in the system, offer great flexibility.  However, they
require deep integration with the programming language and modify
the semantics in non-trivial ways.  This makes fine-grained systems
less suitable in contexts where preserving an existing language
is important.
%
Moreover, the task of adding fine-grained IFC even to a simple language is
non-trivial~\cite{hritcu2013testing}; when considering a real-world language,
such as JavaScript, this challenge becomes almost insurmountable.

In the alternative \textit{coarse-grained} approach to IFC,
a system is divided into relatively coarse computational units,
each with a single label dictating its security policy.
Only
communication between isolated computational units must be tracked.
%A recent system named SWAPI~\cite{swapi} adds IFC to the
%browser by identifying \emph{web workers} as an existing computational
%unit.  Communication between workers is mediated using
%IFC to ensure non-interference, preventing sensitive workers from
%influencing the computation of less sensitive workers.
This coarse-grained approach provides a number of advantages:
(1) minimal changes to an
existing programming language when adding IFC, which (2) allows
the reuse of existing programs.  Furthermore, (3)
reduced runtime overhead because checks need only
be performed at isolation boundaries, compared to fine-grained
systems were the runtime overhead can be impractical~\cite{JSFlow}.
Finally, (4) associating
a single security label with the entire computational unit simplifies
understanding and reasoning about the correctness of the
system, without information flow reasoning about the
technical details of the semantics of the programming language.

A further benefit of the coarse-grained approach
is that a single IFC design and semantics can be generalized
to a large class of programming languages.
Leveraging this insight, we define a form of coarse-grained
dynamic IFC and show how it can be combined with any programming
language for which external effects can be controlled.
At the core of our IFC system is a sandboxing construct that isolates
potentially sensitive computations, allowing them to
communicate only through clearly specified interfaces.
Using an approach by Matthews-Findler~\cite{Matthews:2007:OSM:1190216.1190220}
for combining operational semantics,
we formally state and prove non-interference guarantees that are
independent of the choice of specific target language.

One challenge in the design of our system is that the semantics for the
retrofitted IFC language may not accurately correspond to the optimized
implementation details of a particular target language.
We discuss and partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then non-interference in the abstract specification
carries over to the concrete implementation.

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  information flow control language free of non-IFC constructs.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Section~\ref{sec:retrofit})
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC language that preserves non-interference. (Section~\ref{sec:formal}
  and~\ref{sec:concrete})
  \item We have implemented an IFC system based on these semantics
  for Node.js, and we connect our formalism to another implementation
  based on this work
  for client-side JavaScript in Gecko~\cite{swapi}.
  Furthermore, we outline an implementation for the C programming
  language.
  (Section~\ref{sec:real})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, labeled references, or a
  notion of clearance. (Section~\ref{sec:extensions})
\end{itemize}
